---
title: "The Scale-Ordered Contagion methodology"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The Scale-Ordered Contagion methodology}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 6, fig.height = 4)
library(sochcontagion)
```

## The mechanism

A shock in a source market $s$ releases information that the receiving market
$r$ absorbs through its own exponential filter. The transmitted response is the
convolution of source generation and receiver absorption -- a *bi-exponential*
whose power spectrum is the product of two Lorentzians, one with corner at the
source rate $\alpha_s$ and one at the receiver rate $\alpha_r$:

$$ S(\omega) = A^2\,\frac{\alpha_s^2}{\alpha_s^2+\omega^2}\,
                       \frac{\alpha_r^2}{\alpha_r^2+\omega^2}. $$

The slower market supplies the *binding corner*: it governs the frequency band
in which contagion power concentrates. The spectrum is symmetric in the two
rates -- the basis of the shape-symmetry prediction below.

```{r spectrum}
w <- seq(0, pi, length.out = 200)
plot(w, soch_spectrum(2.0, 0.2, w), type = "l", lwd = 2,
     xlab = expression(omega), ylab = "S(omega)",
     main = "Product-Lorentzian transmission spectrum (fast source, slow receiver)")
```

## From spectrum to wavelet scales

Projecting the spectrum onto MODWT octave bands gives a closed-form
transfer-entropy-by-scale profile, `soch_scale_power()`. The two mixed
directions (fast$\to$slow and slow$\to$fast) have *identical* normalised
profiles, while fast/fast peaks fine and slow/slow peaks coarse:

```{r profiles}
P <- list(
  "fast|fast" = soch_scale_power(2.0, 2.0),
  "fast|slow" = soch_scale_power(2.0, 0.2),
  "slow|fast" = soch_scale_power(0.2, 2.0),
  "slow|slow" = soch_scale_power(0.2, 0.2)
)
plot_scale_profiles(P, normalise = TRUE)
```

The peak frequency obeys $\omega^\star \in [\alpha_\wedge/\sqrt3,\ \alpha_\wedge]$,
and the predicted peak scale is $k^\star \approx \log_2(\pi/\alpha_\wedge)$:

```{r peak}
soch_peak_frequency(2.0, 0.2)   # near the slow rate 0.2
soch_peak_scale(2.0, 0.2)       # a coarse scale
```

## Three falsifiable predictions

* **SOCH-A** -- the peak scale is set by the slower market's rate, so
  emerging-inclusive pairs peak at coarser scales.
* **SOCH-B** -- the *shape* of the scale profile is the same in both
  directions (the spectrum is symmetric in the two rates).
* **SOCH-C** -- the *level* is directionally asymmetric, scaling with
  connectivity and source-shock content.

## Identification

Fitting an observed profile to the closed form recovers the pair's rates by
nonlinear least squares, concentrating out the level. Because the scale power
is symmetric, a single direction identifies only the *unordered* pair
$\{\alpha_\wedge, \alpha_\vee\}$; pooling across pairs that share a market
resolves the ordering and yields market-level rates.

```{r estimator}
truth <- 3 * soch_scale_power(2.0, 0.2)        # known rates + a level
fit <- soch_fit_pair(truth)
c(amin = round(fit$amin, 3), amax = round(fit$amax, 3), R2 = round(fit$R2, 4))
```

See the *Replication* vignette to reproduce the headline G20 results.
