---
title: "Reproducing the G20 results"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reproducing the G20 results}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 6, fig.height = 4)
library(sochcontagion)
```

This vignette reproduces the headline empirical results of the Scale-Ordered
Contagion paper on the bundled G20 equity returns. The light steps run here;
the heavier all-pairs and bootstrap steps are shown with `eval = FALSE` and run
in a few minutes on a laptop.

## Data and the verified directional signature

```{r data}
data(g20_returns)
data(market_groups)
dim(g20_returns)

usa_india <- wqte_profile(g20_returns, "USA", "India", tau = 0.05)
india_usa <- wqte_profile(g20_returns, "India", "USA", tau = 0.05)
round(rbind(`USA->India` = usa_india, `India->USA` = india_usa), 4)
```

The United-States-to-India profile rises through d4 (the fast-source /
slow-receiver signature of SOCH-A) and exceeds the reverse direction in level
(SOCH-C):

```{r validate}
all(diff(usa_india[1:4]) > 0)            # rising d1 -> d4
mean(usa_india) > mean(india_usa)        # level asymmetry
plot_scale_profiles(list(`USA->India` = usa_india, `India->USA` = india_usa))
```

## Test 1 (SOCH-A) and Test 3 (SOCH-C) on the eight-market sample

```{r tests-light}
adv <- c("USA", "UK", "Germany", "Japan")
emg <- c("China", "India", "Brazil", "SouthAfrica")
P <- soch_profiles(g20_returns, c(adv, emg), tau = 0.05)

t1 <- soch_test_ordering(P, emerging = emg)
c(slope = round(t1$slope, 3), p_value = round(t1$p_value, 3))
t1$means                                 # mean peak scale by # emerging

t3 <- soch_test_magnitude(P, advanced = adv, emerging = emg)
c(median_ratio = round(t3$median, 3), frac_gt1 = t3$frac_gt1,
  p_value = round(t3$p_value, 3))
```

## Test 4: endogenous classification

```{r test4}
prof_list <- lapply(names(P), function(k)
  list(i = sub("\\|.*", "", k), j = sub(".*\\|", "", k), wqte = P[[k]]))
rates <- soch_fit_market(prof_list, c(adv, emg))
rates[order(-rates$alpha), ]
plot_market_rates(rates)
```

India and China are recovered as decisively the slowest adapters; the faster
markets cluster near the Nyquist resolution limit and are not separately
identified -- exactly as the identification analysis predicts.

## Test 2 (SOCH-B): the sharp shape-symmetry test

The shape-symmetry test uses a stationary block bootstrap and is heavier; run
it directly:

```{r test2, eval = FALSE}
res <- soch_test_symmetry(g20_returns, c(adv, emg), tau = 0.05, B = 200)
mean(res$holds)            # share of unordered pairs that do not reject SOCH-B
plot_soch_symmetry(res)
```

Or run the entire programme in one call:

```{r pipeline, eval = FALSE}
out <- soch_pipeline(g20_returns, adv, emg, tau = 0.05, symmetry = TRUE)
out$test_ordering$p_value
out$market_rates
```
